### 2012-07-24 Papers/images/Stack.tex
Adding images to paper

\documentclass[10pt,a4paper]{article}
\usepackage[rgb]{xcolor}
\usepackage{tikz}
\usetikzlibrary{calc, positioning}
\usetikzlibrary{fit}
\usetikzlibrary{shapes.geometric}
\usepgflibrary{shapes.geometric}
\usepgflibrary{decorations.shapes}

\begin{document}
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}
\definecolor[named]{fillColor}{rgb}{0.48,0.17,0.26}

\begin{figure}[h!]
 \centering
\definecolor[named]{fillColor}{rgb}{0.48,0.17,0.26}
\begin{tikzpicture}
[block/.style={rectangle,draw=black!50,text width=1.7cm,,align=center,
decoration={shape sep=.1cm}}]

\node[block, minimum height=30mm,minimum width=60mm, fill=gray!20 ] (server)  at (5.3,0) {};
\node[text width=6cm,align=center] (text) at (5.3,2.1)  {Web application server\\(Tomcat)};

\node[block,minimum width=18mm, minimum height=17mm,fill=fillColor!69!white,text=white] (client) {Client};
\node[text width=3cm,align=center] (text) at (0,2.1)  {Web browser\\ (Firefox,Chrome)};

\node[block, minimum width=20mm,minimum height=25mm,fill=fillColor!69!white,text=white] (webserver)
at (3.6,0) {Web server};
\node[block, minimum width=30mm,minimum height=25mm,fill=fillColor!69!white,text=white] (applicationserver)
at (6.4,0) {Application server};

\node[block,fill=black!0 ,minimum width=18mm, minimum height=17mm] (persistence) 		at (8.5,0) {Persistence
provider};

\node[text width=6cm,align=center] (text) at (12,2.1)  {Database\\ (HSQLDB)};
\node[minimum width=1cm, minimum height=2.7cm, cylinder, shape border rotate=90, draw, >=stealth,shape
aspect=.2,fill=fillColor!69!white,text=white] (database) 				at (12,0) {Data source};

\draw [->] (client) to (server);
\draw [<-] (client) to node [above,align=center,midway] {HTTP}  (server);
\draw [<->] (persistence) to node [above,align=center,midway] {JDBC} (database) ;

\end{tikzpicture}
\caption{Starlight web-based application architecture}
\end{figure}

\begin{figure}[h!]
 \centering
\definecolor[named]{fillColor}{rgb}{0.48,0.17,0.26}
\begin{tikzpicture}
[block/.style={rectangle,draw=black!50,text width=1.7cm,,align=center,
decoration={shape sep=.1cm}}]

\node[block, minimum height=20mm,minimum width=60mm, fill=fillColor!69!white,text=white] (apps1)  at (-5.1,0) {};
\node[text width=6cm,align=center,text=white,font=\Large\bfseries\sf ] (text) at (-5.25,0.64)  {Stock Android Apps};
\node[text width=6cm,align=center,text=white,font=\small\sf ] (text) at (-5.25,-0.34)  {Launcher   Phone  Alarm\\Email  Settings  Camera\\Gallery  Calendar ... };

\node[block, minimum height=20mm,minimum width=60mm, fill=fillColor!69!white,text=white ] (apps2)  at (1.1,0) {};
\node[text width=6cm,align=center,text=white, font=\Large\bfseries\sf ] (text) at (0.2,0)  {Custom/Market Apps};

\draw[style=dashed] (-8.5, -1.2) to (3.5,-1.2);

\node[block, minimum height=20mm,minimum width=60mm, fill=fillColor!69!white,text=white] (apps1)  at (-5.1,-2.4) {};

\end{tikzpicture}
\caption{Starlight web-based application architecture}
\end{figure}

\end{document}