### 2012-08-07 Papers/images/Stack.tex
updating files

\documentclass[10pt,a4paper]{article}
\usepackage[rgb]{xcolor}
\usepackage{tikz}
\usetikzlibrary{calc, positioning}
\usetikzlibrary{fit}
\usetikzlibrary{shapes.geometric}
\usepgflibrary{shapes.geometric}
\usepgflibrary{decorations.shapes}
\thispagestyle{empty}
\begin{document}
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}
\definecolor[named]{fillColor}{rgb}{0.48,0.17,0.26}

\begin{figure}[h!]
 \centering
\definecolor[named]{grey}{rgb}{0.9,0.9,0.9}
\definecolor[named]{green}{rgb}{0.297,0.832,0.793}
\definecolor[named]{blue}{rgb}{0.574,0.633,0.926}

\definecolor[named]{fillColor}{rgb}{0.48,0.17,0.26}

\begin{tikzpicture}
[block/.style={rectangle,draw=black,text width=1.7cm,,align=center,
decoration={shape sep=.1cm}}]

\node[block, minimum height=20mm,minimum width=60mm,fill=green] (apps1)  at (-3.25,1.25) {};
\node[text width=6cm,align=center,font=\Large\bfseries\sf ] (text) at (-3.25,1.8)  {Stock Android Apps};
\node[text width=6cm,align=center,font=\small\sf ] (text) at (-3.25,0.9)  {Launcher   Phone  Alarm\\Email  Settings 
Camera\\Gallery  Calendar ... };

\node[block, minimum height=20mm,minimum width=60mm,fill=green] (apps2)  at (3.25,1.25) {};
\node[text width=6cm,align=center, font=\Large\bfseries\sf ] (text) at (3.25,1.25)  {Custom/Market Apps};

\draw[style=dashed] (-7, 0) to (7,0);
\draw[style=dashed] (0, 0) to (0,-3.5);
\node[align=center,font=\small\sf,rotate=90, fill=white] (text) at (0,-1.5)  {Binder};

\node[block, minimum height=30mm,minimum width=60mm,fill=grey] (apps1)  at (-3.25,-1.75) {};
\node[text width=6cm,align=center,font=\Large\bfseries\sf ] (text) at (-3.25,-0.65)  {System Server};
\node[text width=6cm,align=center,font=\small\sf ] (text) at (-3.25,-2)  {Activity Manager    Window Manager\\Package Manager   
Location Manager\\Power Manager Battery Service\\Notification Manager Mount Service\\Clipboard Service  ...};

\node[block, minimum height=30mm,minimum width=28.75mm,fill=grey] (apps1)  at (1.6875,-1.75) {};
\node[text width=6cm,align=center,font=\Large\bfseries\sf ] (text) at (1.6875,-1.75)  {android.*};
\node[block, minimum height=30mm,minimum width=28.75mm,fill=grey] (apps1)  at (4.8125,-1.75) {};
\node[text width=6cm,align=center,font=\Large\bfseries\sf ] (text) at (4.8125,-1.57)  {java.*};
\node[text width=6cm,align=center,font=\small\sf ] (text) at (4.8125,-2)  {(Apache Harmony)};

\node[block, minimum height=10mm,minimum width=125mm,fill=grey] (apps1)  at (0,-4) {};
\node[text width=13cm,align=center,font=\large\bfseries\sf ] (text) at (0,-4)  {Android Runtime/Dalvik/Zygote};

\draw[style=dashed] (-7, -4.75) to (7,-4.75);

\node[block, minimum height=12mm,minimum width=28.75mm,fill=blue] (apps1)  at (1.6875,-5.6) {};
\node[text width=3cm,align=center,font=\large\bfseries\sf ] (text) at (1.6875,-5.6)  {Libs};
\node[block, minimum height=12mm,minimum width=28.75mm,fill=blue] (apps1)  at (4.8125,-5.6) {};
\node[text width=3cm,align=center,font=\large\bfseries\sf ] (text) at (4.8125,-5.6)  {Init/Toolbox};
\node[block, minimum height=12mm,minimum width=28.75mm,fill=blue] (apps1)  at (-1.6875,-5.6) {};
\node[text width=3cm,align=center,font=\large\bfseries\sf ] (text) at (-1.6875,-5.6)  {Native\\Daemons};
\node[block, minimum height=12mm,minimum width=28.75mm,fill=blue] (apps1)  at (-4.8125,-5.6) {};
\node[text width=3cm,align=center,font=\large\bfseries\sf ] (text) at (-4.8125,-5.6)  {Hardware\\Drivers};

\node[block, minimum height=10mm,minimum width=125mm,fill=blue] (apps1)  at (0,-6.95) {};
\node[text width=13cm,align=center,font=\Large\bfseries\sf ] (text) at (0,-6.75)  {Linux Kernel};
\node[text width=6cm,align=center,font=\small\sf ] (text) at (0,-7.25)  {Binder, ashmem, wakelocks,lowmem, ...};

\end{tikzpicture}
\end{figure}

\newpage
\thispagestyle{empty}
\begin{figure}[h!]
 \centering
\definecolor[named]{grey}{rgb}{0.9,0.9,0.9}
\definecolor[named]{green}{rgb}{0.297,0.832,0.793}
\definecolor[named]{blue}{rgb}{0.574,0.633,0.926}

\definecolor[named]{fillColor}{rgb}{0.48,0.17,0.26}

\begin{tikzpicture}
[block/.style={rectangle,draw=black,text width=1.7cm,,align=center,
decoration={shape sep=.1cm}}]

\node[block,minimum height=1cm,minimum width=4.8cm,fill=green,text width=4.8cm,align=center, font=\bfseries\sf ] (application) at
(0cm,1.7cm) {Application Components};

\path [draw,fill=grey ] (-3.8,-4) -- (-3.8,1.6) -- (-2.7,1.6) -- (-2.7,1) -- (2.7,1) -- (2.7,1.6) --  (3.8,1.6) -- (3.8,-4) --
(-3.8,-4);

\node[block, minimum height=1cm,minimum width=2cm,text width=2cm, font=\small\bfseries\sf,fill=white ] (am)  at (-2.5cm,-3.3cm)
{JPF\_Activity Manager};
\node[block, minimum height=1cm,minimum width=2cm,text width=2cm,font=\small\bfseries\sf,fill=white ] (mq)  at (0cm,-3.3cm) {JPF\_Message
Queue};
\node[block, minimum height=1cm,minimum width=2cm,text width=2cm, font=\small\bfseries\sf ,fill=white] (win)  at (2.5cm,-3.3cm)
{JPF\_Window};

\draw[style=dashed,thick] (-5, -2.6) to (5,-2.6);
\node[text width=6cm,align=center, font=\bfseries\sf ] (text1) at (4.3cm,-2.3)  {JPF};
\node[text width=6cm,align=center, font=\bfseries\sf ] (text2) at (4.3cm,-2.9)  {MJI};

 \node[block, minimum height=1cm,minimum width=2.7cm,font=\bfseries\sf,,fill=white] (MQ)  at (-2.25cm,-1.3) {Message Queue};
\node[block, minimum height=1cm,minimum width=7.2cm,align=center, font=\large\bfseries\sf,fill=white ] (apps2)  at (0cm,0.3cm)
{ActivityThread};
% \draw[style=dashed] (-3.8,-1) to (3.8,-1);
 \node[block, minimum height=1cm,minimum width=2.7cm, font=\small\bfseries\sf,fill=white ] (loop)  at (2.25cm,-1.3cm) {Looper};
%  \node[block, minimum height=1cm,minimum width=2cm,text width=2cm, font=\small\bfseries\sf ] (apps2)  at (2.5cm,-1.9cm) {};

\node[minimum width=1cm,text width=1cm, minimum height=1.4cm, cylinder, shape border rotate=90, draw, >=stealth,shape
aspect=.2,fill=green, align=center] (database) 				at (0,-5) {Input\\ Script};

\draw [->, thick] (win) to [ out=30, in= -10] (application)  ;
\node[text width=3cm,align=center, font=\small\bfseries\sf ] (text) at (5,-1)  {\$button.onClick()};

\draw [->,thick] (database) to (mq);

\draw [->,thick] (mq) to (am);
\draw [->,thick] (mq) to (win);
\draw [->, thick] (am) to  [ out=180, in=180] (MQ);
\draw [->, thick] (MQ) to (mq);
\draw [->, thick] (loop) to (MQ);
\draw [->, thick] (loop) to (apps2);
\node[text width=3cm,align=center, font=\small\bfseries\sf ] (text) at (1.8cm,-0.5cm)  {handleLaunchActivity(...)};
\node[text width=3cm,align=center, font=\small\bfseries\sf ] (text) at (-4,-2.3)  {scheduleLaunchActivity(...)};

\node[text width=3cm,align=center, font=\small\bfseries\sf ] (text) at (0.5, -2.3)  {processScriptAction()};
\node[text width=3cm,align=center, font=\small\bfseries\sf ] (text) at (0, -1.1)  {getNext()};

\draw [->,thick] (apps2) to (application);

\end{tikzpicture}
\end{figure}

\end{document}